---
output: github_document
---

# prepost


<!-- badges: start -->  [![R-CMD-check](https://github.com/mattblackwell/prepost/workflows/R-CMD-check/badge.svg)](https://github.com/mattblackwell/prepost/actions)
<!-- badges: end -->


## Overview

This package contains code and sample data to implement the non-parametric bounds and Bayesian methods for assessing priming and post-treatment bias in experimental studies under various assumptions.  


To get started, please see the article that developed these methods:

* Blackwell, Matthew et al (2025). Priming Bias Versus Post-Treatment Bias in Experimental Designs. *Political Analysis* Published online 2025:1-17. doi:10.1017/pan.2025.3. ([journal version](http://doi.org/10.1017/pan.2025.3), [arXiv preprint](https://arxiv.org/pdf/2306.01211))


## Installation

```{r install, eval = FALSE}
## Install developer version
## install.packages("devtools")
devtools::install_github("mattblackwell/prepost", build_vignettes = TRUE)
```


## Usage

Both the nonparametric and Bayesian estimators all have prefixes that indicate what type of experimental design being used.  

- `pre_` functions can analyze data from a **pre-test design** where the moderator is measured pre-treatment. 
- `post_` functions can analyze data from a **post-test design** where the moderator is measured post-treatment.
- `prepost_` functions can analyze data from a **random placement design**, in which the moderator is randomly assigned to be measured before or after treatment.

Most functions can be specified with a formula to identify the outcome and treatment and another one-sided formula for the moderator: 

```{r}
library(prepost)
data(delponte)
out <- pre_bounds(
  formula = angry_bin ~ t_commonality,
   data = delponte,
  moderator = ~ itaid_bin
)
out
```

## Citation

